---
title: "homework 3"
author: "lauren franks, lef2147"
date: "10/3/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 9,
  fig.asp = .6,
  out.width = "90%"
)

library(dplyr)
library(ggplot2)
library(viridis)
library(tidyverse)
```

Problem 1
Loading Data
```{r}
library(p8105.datasets)
data("instacart")
```

To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations
```{r, include=FALSE}
nrow(instacart)
ncol(instacart)
skimr::skim(instacart)
str(instacart)
```

```{r, warning=FALSE}
ggplot(instacart, aes(days_since_prior_order)) + 
  geom_histogram() + 
  labs(title = "Histogram of days since prior order",
       x = "Number of days since prior order",
       y = "Frequency")

instacart %>% filter(days_since_prior_order < 30) %>% 
  ggplot(aes(days_since_prior_order)) +
  geom_histogram() + 
  labs(title = "Histogram of days 1 to 29 since prior order",
       x = "Number of days since prior order",
       y = "Frequency")

instacart %>% count(days_since_prior_order == 30)
```
There are `r ncol(instacart)` variables in the instacart dataset, and `r nrow(instacart)` observations. There are 4 character variables, and 11 integer variables. Investigating the variable days_since_prior_order, based on the histogram we can see that it is right skewed, with a high proportion of values at 30 days. Removing the observations that = "30", we can see that there are several peaks across the variable. The peaks are at 7, 14, 21 and 28 days, with the highest peak at 7 days. This may suggest that people generally reorder from instacart on the same day of the week that they made their previous order. The high proportion of observations categorized as 30 suggests that anyone who orders past 30 days gets grouped into "30" in this dataset.

How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% group_by(aisle_id) %>% 
  summarise(n_obs = n(), n_unique = n_distinct(aisle_id)) %>% 
  arrange(desc(n_obs))
```
Aisle 83 has the most orders, and there are 134 aisles in total. 


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10,000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>% group_by(aisle_id) %>% 
  tally() %>% 
  subset(n > 10000) %>% 
  ggplot(aes(aisle_id, n, label = aisle_id)) + 
  geom_point() + labs(
    title = "Number of Items ordered by Aisle",
    x = "Aisle Number",
    y = "Number of Items Ordered"
  ) + scale_x_continuous(
    breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140)
  ) + scale_y_continuous(
    breaks = c(20000, 40000, 60000, 80000, 100000, 120000, 140000)
  )
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  group_by(aisle) %>% 
  top_n(3) %>% 
  as.tibble() %>% 
  select(-aisle) %>% 
  knitr::kable(col.names = c("Product Name", "Number of items ordered"))
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
```{r}
instacart %>% select(order_dow, order_hour_of_day, product_name) %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(order_dow, order_hour_of_day, product_name) %>% 
  summarise(n_obs = n()) %>% 
  group_by(product_name, order_dow) %>% 
  filter(n_obs == max(n_obs)) %>% 
  summarise(mean_hour = mean(n_obs)) %>% 
  tidyr::pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% knitr::kable(col.names = c("Product Name", "Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"))
```


Question 2
Loading Data
```{r}
data("brfss_smart2010")

str(brfss_smart2010)
```

Data Cleaning
 - format the data to use appropriate variable names
 - focus on the “Overall Health” topic
 - include only responses from “Excellent” to “Poor”
 - organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```{r}
brfss = janitor::clean_names(brfss_smart2010) %>% 
  rename(
    state_abb = locationabbr,
    county = locationdesc
  ) %>% 
  filter(topic == "Overall Health") %>% 
  mutate(
    response = as.factor(response),
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE)
    ) %>% 
  arrange(response)
```

Questions:
In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss %>% filter(year == 2002) %>% 
  group_by(county, state_abb) %>% 
  summarise(n_obs = 1) %>% 
  group_by(state_abb) %>% 
  summarise(sum_loc = sum(n_obs)) %>% 
  subset(sum_loc >= 7)

brfss %>% filter(year == 2010) %>% 
  group_by(county, state_abb) %>% 
  summarise(n_obs = 1) %>% 
  group_by(state_abb) %>% 
  summarise(sum_loc = sum(n_obs)) %>% 
  subset(sum_loc >= 7)
```
In 2002, there were 6 states observed at 7 or more locations (CT, FL, MA, NC, NJ, and PA), and in 2010 there were 14 states (CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX and WA). 

Construct a dataset that is limited to Excellent responses, and contains year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss_exc = brfss %>% filter(response == "Excellent") %>% 
  select(year, state_abb, county, data_value) %>% 
  group_by(state_abb, year) %>% 
  summarise(
    avg_data_value = mean(data_value, na.rm = TRUE)
  )

brfss_exc %>% ggplot(aes(year, avg_data_value, group = state_abb)) + 
  geom_line() + labs(
    title = "Average Data Value By State",
    subtitle = "For Response = Excellent",
    x = "Year",
    y = "Average Data Value"
  )
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss %>% filter(year == c(2006, 2010), state_abb == "NY") %>% 
  group_by(county, response) %>% 
  select(year, county, response, data_value) %>% 
  ggplot(aes(response, data_value, group = county, color = county)) + 
  geom_point() +
  facet_grid(~year) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```





