---
title: "homework 3"
author: "lauren franks, lef2147"
date: "10/3/2019"
output: github_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(viridis)
library(tidyverse)
```

Problem 1
Loading Data
```{r}
library(p8105.datasets)
data("instacart") 
```

To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations
```{r, include=FALSE}
nrow(instacart)
ncol(instacart)
skimr::skim(instacart)
str(instacart)
```

```{r, warning=FALSE}
ggplot(instacart, aes(days_since_prior_order)) + 
  geom_histogram() + 
  labs(title = "Histogram of days since prior order",
       x = "Number of days since prior order",
       y = "Frequency")

instacart %>% filter(days_since_prior_order < 30) %>% 
  ggplot(aes(days_since_prior_order)) +
  geom_histogram() + 
  labs(title = "Histogram of days 1 to 29 since prior order",
       x = "Number of days since prior order",
       y = "Frequency")

instacart %>% count(days_since_prior_order == 30)

total_prod_ordered = instacart %>% group_by(order_id) %>% 
  summarise(
    total_products_ordered = n()
  )
```
There are `r ncol(instacart)` variables in the instacart dataset, and `r nrow(instacart)` observations. There are 4 character variables, and 11 integer variables. The mean number of items per order is `r mean(pull(total_prod_ordered, total_products_ordered))`. Investigating the variable days_since_prior_order, based on the histogram we can see that it is right skewed, with a high proportion of values at 30 days. Removing the observations that = "30", we can see that there are several peaks across the variable. The peaks are at 7, 14, 21 and 28 days, with the highest peak at 7 days. This may suggest that people generally reorder from instacart on the same day of the week that they made their previous order. The high proportion of observations categorized as 30 suggests that anyone who orders past 30 days gets grouped into "30" in this dataset. 

How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% group_by(aisle_id) %>% 
  summarise(n_obs = n(), n_unique = n_distinct(aisle_id)) %>% 
  arrange(desc(n_obs))
```
Aisle 83 has the most orders, and there are 134 aisles in total. 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10,000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>% group_by(aisle_id) %>% 
  tally() %>% 
  subset(n > 10000) %>% 
  ggplot(aes(aisle_id, n, label = aisle_id)) + 
  geom_point() + labs(
    title = "Number of Items ordered by Aisle",
    subtitle = "Aisles with greater than 10,000 items ordered",
    x = "Aisle Number",
    y = "Number of Items Ordered"
  ) + scale_x_continuous(
    breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140)
  ) + scale_y_continuous(
    breaks = c(20000, 40000, 60000, 80000, 100000, 120000, 140000)
  )
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  group_by(aisle) %>% 
  top_n(3) %>% 
  as.tibble() %>% formattable::formattable(col.names = c("Aisle", "Product Name", "Number of items ordered"))
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
```{r}
instacart %>% select(order_dow, order_hour_of_day, product_name) %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(order_dow, order_hour_of_day, product_name) %>% 
  summarise(n_obs = n()) %>% 
  group_by(product_name, order_dow) %>% 
  filter(n_obs == max(n_obs)) %>% 
  summarise(mean_hour = mean(n_obs)) %>% 
  tidyr::pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% knitr::kable(col.names = c("Product Name", "Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"))
```


Question 2
Loading Data
```{r, include=FALSE}
data("brfss_smart2010")

str(brfss_smart2010)
```

Data Cleaning
 - format the data to use appropriate variable names
 - focus on the “Overall Health” topic
 - include only responses from “Excellent” to “Poor”
 - organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```{r}
brfss = janitor::clean_names(brfss_smart2010) %>% 
  rename(
    state_abb = locationabbr,
    county = locationdesc
  ) %>% 
  filter(topic == "Overall Health") %>% 
  mutate(
    response = as.factor(response),
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE)
    ) %>% 
  arrange(response) %>% select(-c(class, topic, question))
```

Questions:
In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss %>% filter(year == 2002) %>% 
  group_by(county, state_abb) %>% 
  summarise(n_obs = 1) %>% 
  group_by(state_abb) %>% 
  summarise(sum_loc = sum(n_obs)) %>% 
  subset(sum_loc >= 7)

brfss %>% filter(year == 2010) %>% 
  group_by(county, state_abb) %>% 
  summarise(n_obs = 1) %>% 
  group_by(state_abb) %>% 
  summarise(sum_loc = sum(n_obs)) %>% 
  subset(sum_loc >= 7)
```
In 2002, there were 6 states observed at 7 or more locations (CT, FL, MA, NC, NJ, and PA), and in 2010 there were 14 states (CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX and WA). 

Construct a dataset that is limited to Excellent responses, and contains year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss_exc = brfss %>% filter(response == "Excellent") %>% 
  select(year, state_abb, county, data_value) %>% 
  group_by(state_abb, year) %>% 
  summarise(
    avg_data_value = mean(data_value, na.rm = TRUE)
  )

brfss_exc %>% ggplot(aes(year, avg_data_value, group = state_abb)) + 
  geom_line() +
  labs(
    title = "Average Data Value By State",
    subtitle = "For Response = Excellent, from years 2002 - 2010",
    x = "Year",
    y = "Average Data Value"
  ) 
```
This plot shows the trend of the average data value across all counties per state from 2002 to 2010. Because of the density of datapoints in this plot, it is difficult to take away information from it. Labels make it too jumbled to read. You can see there is a slight, general downward trend; in 2002 there were several states with average data values close to 30, whereas in 2010 most states were below 27.5. 

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss %>% filter(year == c(2006, 2010), state_abb == "NY") %>% 
  group_by(county, response) %>% 
  select(year, county, response, data_value) %>% 
  ggplot(aes(response, data_value, group = county, color = county)) + 
  geom_point() +
  facet_grid(~year) +
  labs(
    x = "Response",
    y = "Data Value"
  ) + 
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )

brfss %>% filter(year == c(2006, 2010), state_abb == "NY") %>% 
  group_by(county, response) %>% 
  select(year, county, response, data_value) %>% 
  group_by(county)
```
In both 2006 and 2010, "good" and "very good" responses had on average higher data values than "excellent" responses. There is a smaller spread in response "very good" in 2010 compared to "very good" in 2006, and poor responses have a greater spread in 2010 compared to poor reponses in 2006. Both graphs show the same overall trend with a general increase from Poor -- Very Good and a slight decrease for excellent responses. Counties who were interview in 2006 were not neccessarily interviewed in 2010, which could introduce some data discrepancies when comparing the data vales between the two years. 

Question 3:
Loading Data
Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r}
accel_data = read_csv("./accel_data.csv")
```

Tidying Data
```{r}
accel_data = accel_data %>% janitor::clean_names() %>% 
  mutate(
    weekend = ifelse(day == c("Saturday", "Sunday"), "Weekend", "Weekday"),
    day_ordered = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"), ordered = TRUE)
  ) %>% 
  select(week, day_id, day, day_ordered, weekend, activity_1:activity_1440)
```
The dataset consists of 1444 variables, a week variable ranging from 1-5, a day ID variable ranging from 1-35, a day variable indicating which day of the week it is, a weekend variable indicate if it is a weekend observation or weekday observation, and 1440 variables for each minute of the day that measures activity count. There are 35 observations for 35 days. 

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
accel_data_total = accel_data %>% 
  mutate(
    total_activity = rowSums(accel_data[, grep("activity_", names(accel_data))])
  ) %>% 
  select(week, day_ordered, total_activity) %>% 
  arrange(day_ordered) %>% 
  group_by(week, day_ordered) %>% 
  pivot_wider(
    names_from = "day_ordered",
    values_from = "total_activity"
  ) %>% view()

accel_data %>% 
  mutate(
    total_activity = rowSums(accel_data[, grep("activity_", names(accel_data))])
  ) %>% 
  select(week, day_ordered, total_activity) %>% 
  group_by(week, day_ordered) %>% 
  ggplot(aes(day_ordered, total_activity, group = week, color = week)) + 
  geom_line() +
  viridis::scale_color_viridis(
    name = "week",
    discrete = FALSE
  )
```
Saturday for week 4 and 5 both have the value 1440, which means that for each activity minute, there was a 1.0 entered. This might suggest an error by the accelerometer, or that the subject was not wearing it. I created a line plot to better visualize any trends. There is a very large spread on Fridays, Saturdays, Sundays, and Mondays, and a very narrow spread on Tuesdays, Wednesdays, and Thursdays. This might suggest that the activities that this person is doing on those days are similar, whereas they tend to vary more Friday -- Monday. 

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
accel_long = accel_data %>% select(week, day, activity_1:activity_1440) %>% 
  pivot_longer(
    cols = starts_with("activity_"),
    names_to = "activity_minute",
    names_prefix = "activity_",
    values_to = "activity"
  ) 

accel_long %>% 
  ggplot(aes(activity_minute, activity, group = day, color = day)) +
  geom_line(alpha = 0.7) +
  viridis::scale_color_viridis(
    name = "day",
    discrete = TRUE
  )

accel_long %>% filter(week == 1) %>% 
  ggplot(aes(activity_minute, activity, group = day, color = day)) +
  geom_line(alpha = 0.7) +
  viridis::scale_color_viridis(
    name = "day",
    discrete = TRUE
  )
```
Based on this chart with all 5 weeks included, it is a bit difficult to glean information from it, however we are able to distinguish that Wednesday has a very low level energy expenditure across the board. We can also see that Fridays (or whats in purple), has spurts of high energy expenditure. Because there is so many datapoints, I limited the output to only week 1 to make it easier to read. The output is provided. Here you are able to see that Wednesay and Tuesday have very low energy expenditure, and Friday has spurts of very high energy expenditure, consistant with the observations from the original dataset. 
